---
title: "States & Beer"
author: "Joshua Yi"
date: "October 15, 2018"
output:
  html_document:
    css: custom.css
    df_print: paged
    fig_caption: yes
    keep_md: yes
---
```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE,echo=FALSE,comment = NA)
#Loading in the libraries
library(dplyr)
library(kableExtra)
library(formattable)
library(ggplot2)
library(maps)
library(plyr)
library(openintro)
library(stringr)
library(gridExtra)
library(tibble)
library(plotly)
```

# **Introduction**
The analysis we have done will give you a summary level overview of the U.S. Craft Beers and Breweries dataset, that was supplied to us by the CFO and CEO of Budweiser. The overview will contain breweries per state, summary of report on alcohol and IBU content, and correlation between bitterness and alcohol content. After the presentation, you will walk away with a better understanding on areas to focus and the types of beers that are favorable for the consumers.

### *Breweries per state?*

>##### R-code Explanation:
 - Reads in the datasets into a variable
 - Counts brewery by State and sorts in descending order
 - Mutate the table to create a barplot table
 - Create new column to translate fully spelled out states into abbreviations
 - Merge the brewery count by state with the map_data from R a table. This is needed to create a heatmap.

>##### Analysis Explanation:
The provided heat map and bar plot will show you that Colorado has 47 breweries, the highest in the US. While states like Washington D.C., North Dakota, South Dakota and West Virginia only have 1 brewery.

```{r brewery}
# read in csv file 
df <- read.csv('E:\\Documents\\School\\MSDS 6306\\Case Study\\Breweries.csv',header =TRUE, sep=",")
# clean up white spaces
df["State"] <- str_trim(df$State,"left")
# Counting breweries by state
df.ct <- dplyr::count(df,df$State)
# sorting by descending
df.ct <- df.ct[order(-df.ct$n),]
# colnames
colnames(df.ct) = c("State","CountofBreweries")
# creating barplot
table.graph <- df.ct %>%
  mutate(
    # barplot color
    CountofBreweries = color_bar("lightblue")(CountofBreweries)
  ) %>%
    select(State, everything()) %>%
    # setting the table as kable format and giving caption
    kable("html", escape = F,caption = "Table 1: Barplot Brewery Count in Each State") %>%
    # setting kable table style
    kable_styling(bootstrap_options = c("hover"),font_size = 15,full_width = F) %>%
    # changing the width of table column
    column_spec(2, width = "7cm") %>%
    # changing scroll width and height
    scroll_box(width = "400px", height = "450px")

# calling map_data function to setup heatmap
all_states <- map_data("state")
# changing full statname to abbreviated name
all_states["State"] <- state2abbr(all_states$region)
# merging brewery data with map_data for heatmap
Total <- merge(all_states,df.ct, by.x = c("State"), by.y = c("State"))

#ggplot to variable
p <- ggplot()
# customizing the heatmap
p <- p + geom_polygon(data=Total, aes(x=long, y=lat, group = State, fill=Total$CountofBreweries),colour="white"
      ) + scale_fill_continuous(low = "lightgreen", high = "darkgreen", guide="colorbar")
P1 <- p  + theme_bw()  + labs(fill = "Brewery Count",title = "Number of Breweries by State", x="", y="",caption = "Figure 1: Heatmap of Breweries in each State")
p2 <- P1 + scale_y_continuous(breaks=c()) + scale_x_continuous(breaks=c()) + theme(panel.border = element_blank())

# outputs
p2
table.graph
  
```

### *Merging the two datasets. Print the first 6 observations and the last six observations to check the merged file.*
>##### R-code Explanation:
  - Read in the beer dataset
  - Merge the beer data with the brewery data
  - Create a kable table to output the first and last 6 observations
  
>##### Analysis Explanation:
In order to get a better idea and analysis on the data we to have merge the U.S. Craft Beers with the Breweries dataset. This allows us to see the breweries in each state, as well as the types of beers, its alcohol content and IBU that each brewery produces. Tables 2 and 3 are an output as a q/a check on the merger of the data. As you can see it lists the brewery in the State along with the details on the beers that it produces.

```{r Merge}
# reading in beer csv file
beer <- read.csv('E:\\Documents\\School\\MSDS 6306\\Case Study\\Beers.csv',header =TRUE, sep=",")
# merging the beer data with brewery data
m.bb <- merge(df,beer, by.x = c("Brew_ID"), by.y = c("Brewery_id"))
# renaming columns
colnames(m.bb)[colnames(m.bb) == 'Name.x'] <- 'Brewery Name'
colnames(m.bb)[colnames(m.bb) == 'Name.y'] <- 'Beer Name'

# getting the first 6 obs and placing into custom kable table
head.bb <- data.frame(head(m.bb),row.names = NULL) %>%
    kable("html", escape = F,caption = "Table 2: First 6 Observations") %>%
    kable_styling(bootstrap_options = "striped",font_size = 13,full_width = T,position = "float_left") %>%
    column_spec(4, background = "lightgreen") %>%
    column_spec(4, border_right = T,border_left = T)
# output
head.bb

# getting the last 6 obs and placing into custom kable table
tail.bb <- kable(data.frame(tail(m.bb),row.names = NULL),"html", escape = F,caption = "Table 3: Last 6 Observations") %>%
    kable_styling(bootstrap_options = c("striped","hover"),font_size = 13,position = "float_left") %>%
    column_spec(4, background = "lightgreen") %>%
    column_spec(4, border_right = T,border_left = T)

# output
tail.bb

```
### *Reporting out NA's in each column*
>##### R-code Explanation:
  - Calculating the number of NA's in each column
  - Creating bar chart for each NA count

>##### Analysis Explanation:
Further analysis of the data showed that there are missing values in the U.S. craft beer dataset. The below chart shows you the count from each relevant columns. Analysis is only done with the values on-hand.

```{r NA}
# creating count of na's in the brewid column
brewid.na <- cbind(ddply(m.bb,.(),summarise,Count = sum(is.na(m.bb$Brew_ID))),Column = c("BrewID"))
# creating count of na's in the brewery name column
brewnm.na <- cbind(ddply(m.bb,.(),summarise,Count = sum(is.na(m.bb$`Brewery Name`))),Column = c("BrewName"))
# creating count of na's in the city column
city.na <- cbind(ddply(m.bb,.(),summarise,Count = sum(is.na(m.bb$City))),Column = c("City"))
# creating count of na's in the state column
state.na <- cbind(ddply(m.bb,.(),summarise,Count = sum(is.na(m.bb$State))),Column = c("State"))
# creating count of na's in the beer name column
beernm.na <- cbind(ddply(m.bb,.(),summarise,Count = sum(is.na(m.bb$`Beer Name`))),Column = c("Beername"))
# creating count of na's in the beer id column
beerid.na <- cbind(ddply(m.bb,.(),summarise,Count = sum(is.na(m.bb$Beer_ID))),Column = c("BeerID"))
# creating count of na's in the abv column
abv.na <- cbind(ddply(m.bb,.(),summarise,Count = sum(is.na(m.bb$ABV))),Column = c("ABV"))
# creating count of na's in the ibu column
ibu.na <- cbind(ddply(m.bb,.(),summarise,Count = sum(is.na(m.bb$IBU))),Column = c("IBU"))
# creating count of na's in the style column
style.na <- cbind(ddply(m.bb,.(),summarise,Count = sum(is.na(m.bb$Style))),Column = c("Style"))
# creating count of na's in the ounces column
oz.na <- cbind(ddply(m.bb,.(),summarise,Count = sum(is.na(m.bb$Ounces))),Column = c("Ounces"))

# rowbinding all the na variables into one table
na.summary <- rbind(brewid.na,brewnm.na,city.na,state.na,beernm.na,beerid.na,abv.na,ibu.na,style.na,oz.na)
# creating bar chart on na's
plot_ly(na.summary, x = ~Column, y = ~Count, name = ~Column, type = "bar") %>%
    layout(margin = list(l = 2),title = 'NA Count by Column',xaxis = list(title = "Figure 2: NA Count"))


```

### *Summary Analysis*
>##### R-code Explanation:
  - Calculate overall summary statistics on the original dataset; mean, median, max, min and 75th quantile
  - Create interactive line graph

>##### Analysis Explanation:
It's important to understand this because if we just look at the maximum ABV and IBU levels. The perception is that Colorado only produce beers with high alcohol content or Oregon only produce beers that are really bitter. It's the combination of the bar graph and Figure 3 that you start see that Colorado doesn't exclusively produces beers with high alcohol content, but instead they make a couple of beers with high ABV. That's why you see the mean of ABV is being pulled to the higher levels, since it is not resistant to outliers. 
  
```{r report}
# creating columns "Stat" that gives different statistics on the ABV and "Stat_type" which lists out the statistic type
# creating mean stats
Mean.abv <- cbind(ddply(m.bb,.(State),summarise,Stat = mean(ABV,na.rm = T)),Stat_type = c("Mean"))
# creating median stats
Median.abv <- cbind(ddply(m.bb,.(State),summarise,Stat = median(ABV,na.rm = T)),Stat_type = c("Median"))
# creating 75th quantile
Quantile.abv <- cbind(ddply(m.bb,.(State),summarise,Stat = quantile(ABV,c(.75),na.rm = T)),Stat_type = c("75th_Quant"))
# creating min stats
Min.abv <- cbind(ddply(m.bb,.(State),summarise,Stat = min(ABV,na.rm = T)),Stat_type = c("Min"))
# creating max stats
Max.abv <- cbind(ddply(m.bb,.(State),summarise,Stat = max(ABV,na.rm = T)),Stat_type = c("Max"))

# rowbinding all the summary stats into one table
sum.stat <- rbind(Mean.abv,Median.abv,Quantile.abv,Min.abv,Max.abv)

# creating linechart fo the sum.stat table
sum.graph <- plot_ly(sum.stat, x = ~State, y = ~Stat, name = ~Stat_type, type = "scatter", mode = "lines",linetype = ~Stat_type) %>%
    layout(title = 'Summary Statistics for ABV',xaxis = list(title = "Figure 3: Summary Statistics"))
sum.graph
# html output file
htmlwidgets::saveWidget(as_widget(sum.graph),"Sum_graph.html")

```

### *States that have the maximum alcoholic beer or the most bitter beer.*
>##### R-code Explanation:
  - Calculating the median ABV and IBU for each state and ignoring the NA's
  - Creating a interactive barchart for the forementioned calculation for each state

>##### Analysis Explanation:
It's important to look at the maximum, means and medians of the data, so you are able to understand what the consumer wants are for each state. If you look at the graph below you will see the median results for alcohol and bitterness content by state. Here you will start see the different combination of results from ABV to IBU. For example, Washington D.C. has the highest median value of ABV, but Maine has the highest median value for IBU. Is there a correlation?

```{r Max_barchart}
# getting the row which contains the maximum ABV
topabv <- data.frame(m.bb[which.max(m.bb$ABV),],row.names = NULL)
# getting the row which contains the maximum IBU
topibu <- data.frame(m.bb[which.max(m.bb$IBU),],row.names = NULL)

# creating barchart for the max IBU and ABV and placing them into the same grid
gridExtra::grid.arrange(
# barchart for max ABV
ggplot(topabv, aes(x = State, y = ABV, fill = 24))+ geom_bar(stat="identity")+ theme(legend.position="none",plot.title = element_text(size = 15, hjust=.5, face = "bold"))+ggtitle("Max ABV")+labs(caption = "Figure 4: State with max alcohol content")
# barchart for max IBU
,ggplot(topibu, aes(x = State, y = IBU, fill = 24))+ geom_bar(stat="identity")+ theme(legend.position="none",plot.title = element_text(size = 15, hjust=.5, face = "bold"))+ggtitle("Max IBU")+labs(caption = "Figure 5: State with max international bitterness unit")
,nrow = 1)

```


### *Computing the Median Alcohol Content (ABV) and International Bitterness Unit (IBU) for each state.*
>##### R-code Explanation:
  - Calculating the median ABV and IBU for each state and ignoring the NA's
  - Creating a interactive barchart for the forementioned calculation for each state

>##### Analysis Explanation:
It's important to look at the maximum, means and medians of the data, so you are able to understand what the consumer wants are for each state. If you look at the graph below you will see the median results for alcohol and bitterness content by state. Here you will start see the different combination of results from ABV to IBU. For example, Washington D.C. has the highest median value of ABV, but Maine has the highest median value for IBU. Is there a correlation?

```{r Barchart}
# calculating the median abv and ibu for each state, setting into variable
median_m.bb <- ddply(m.bb, .(State), summarize,Median_ABV = median(ABV, na.rm = TRUE), Median_IBU = median(IBU, na.rm = TRUE))
# creating barchart for the median ABV
plot_ly(median_m.bb, x = ~reorder(State,-Median_ABV), y = ~Median_ABV, name = ~State, type = "bar") %>%
    layout(margin = list(l = 2),title = 'Median Alcohol Content',xaxis = list(title = "State   Figure 6: Median ABV" ,tickangle = -90),showlegend = FALSE)
# creating barchart for median IBU
plot_ly(median_m.bb, x = ~reorder(State,-Median_IBU), y = ~Median_IBU, name = ~State, type = "bar") %>%
    layout(margin = list(l = 2),title = 'Median International Bitterness Unit',xaxis = list(title = "State   Figure 7: Median IBU" ,tickangle = -90),showlegend = FALSE)

```

### *Is there an apparent relationship between the bitterness of the beer and its alcoholic content?*
>##### R-code Explanation:
  - Create a interacitve scatterplot with the ABV and IBU data

>##### Analysis Explanation:
Figure 8 is where you see the correlation of a high ABV and its IBU counterpart. It's visible that as the alcohol content level gets higher, so does the IBU level. There are some outliers where the highest ABV at .125 does not have the highest IBU. The importance here is to look at where the clustering is happening. This is your indication to produce a beer that may have a favorable outcome to the majority of the consumers. 

```{r Scatter_plot}
# removing the NA's from the ABV and IBU columns
m.bb_scatter <- subset(x=m.bb,!(is.na(m.bb$ABV)) & !(is.na(m.bb$IBU)))
# creating scatter plot from the m.bb_scatter variable for ABV and IBU
plot_ly(m.bb_scatter, x = ~ABV, y = ~IBU, mode = "markers", type = "scatter", color = ~ABV, size = ~ABV) %>%
    layout(title = 'Relationship between IBU & ABV',xaxis = list(title = "ABV   Figure 8: Scatterplot Relationship"),showlegend = FALSE)

# getting count of beer by state and adding in "type" column
ct.beer <- cbind(ddply(m.bb,.(State),summarise,CT = n()),Type = c("Beer"))
# getting count of breweries by state and adding in "type" column
ct.brew <- cbind(ddply(df,.(State),summarise,CT = n()),Type = c("Brew"))
# rowbinding the two variables to create one table
brew.beer <- rbind(ct.beer, ct.brew)
# creating barchart of the brew.beer table to show visual counts of beer and brew by state.
bar.plot <- plot_ly(brew.beer, x = ~reorder(State,-CT), y = ~CT, name = ~Type, type = "bar") %>%
    layout(title = 'Beer and Brewery Count',xaxis = list(title = "State   Figure 9: Beers and Breweries"))
bar.plot
# html output file
htmlwidgets::saveWidget(as_widget(bar.plot),"bar_plot.html")

```

### *Conclusion*
The best action to take when trying to compete with breweries is to understand the amount of different beers and breweries are in each state. Figure 9 gives you that view. Hit the states with a low brewery and beer type count. Also, look at the clustering in Figure 8 to produce the most favorable beer. For instance, most of the clustering happens to be around ABV levels .04 to .06 and IBU levels of 20 to 40. Hitting the areas that is most acceptable to a newer beer and/or manufacturing a beer with the best ABV to IBU ratio is the best strategy.