---
title: "LiveSession5"
author: "Joshua Yi"
date: "September 27, 2018"
output: 
  html_document:
    keep_md: TRUE
    df_print: paged
---

# Q1a Data Munging
  * First, import the .txt file into R so you can process it. Keep in mind this is not a CSV file. You might have to open the file to see what you're dealing with. Assign the resulting data frame to an object, df, that consists of three columns with human-readable column names for each.

```{r baby_name, echo=TRUE}
# reading from txt file
df <- read.delim("E://Documents//School//MSDS 6306//Unit 5//yob2016.txt",header = FALSE, sep = ";")
# adding column names
colnames(df) = c("FirstName","Gender","Amount")
# output
df
```

# Q1b Data Munging
  * Display the summary and structure of df

```{r baby_name2, echo=TRUE}
# summary
summary(df)
# structure
class(df)

```

# Q1c Data Munging
  * Your client tells you that there is a problem with the raw file.  One name was entered twice and misspelled.  The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y's at the end of the name. Write an R command to figure out which name it is and display it.

```{r baby_name3, echo=TRUE}
# getting row number for the mistake in name
grep("yyy",df$FirstName)
# displyaing the name
df$FirstName[c(212)]
```

# Q1d Data Munging
  * Upon finding the misspelled name, please remove this particular observation, as the client says it's redundant. Save the remaining dataset as an object: y2016 

```{r baby_name4, echo=TRUE}
# removing the mistake
y2016 <- df[-212,]
# displaying
y2016

```


# Q2a Data Merging
  * Like 1a, please import the .txt file into R. Look at the file before you do. You might have to change some options to import it properly. Again, please give the dataframe human-readable column names. Assign the dataframe to y2015.  

```{r merger, echo=TRUE}
# reading from txt file
y2015 <- read.delim("E://Documents//School//MSDS 6306//Unit 5//yob2015.txt",header = FALSE, sep = ",")
# adding column names
colnames(y2015) = c("FirstName","Gender","Amount")
# output
y2015
```

# Q2b Data Merging
  * Display the last ten rows in the dataframe.  Describe something you find interesting about these 10 rows.

```{r merger2, echo=TRUE}
tail(y2015, 10)
"The last ten rows are all males with only 5 people who used that name."

```

# Q2c Data Merging
  * Merge y2016 and y2015 by your Name column; assign it to final.  The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.
```{r merger3, echo=TRUE}
# merging y2016 and y2015
final <- merge(y2016,y2015, by.x = c("FirstName","Gender"), by.y = c("FirstName","Gender"))
colnames(final) = c("FirstName","Gender","Amount.y16","Amount.y15")
# output
final
```
# Q3a Data Summary
  * Create a new column called "Total" in final that adds the amount of children in 2015 and 2016 together.  In those two years combined, how many people were given popular names?
```{r summary, echo=TRUE}
# creating the total column
final["Total"] <- final$Amount.y16+final$Amount.y15
# output
final
```

# Q3b Data Summary
  * Sort the data by Total. What are the top 10 most popular names?
```{r summary2, echo=TRUE}
# sorting the data by descending
sort.final <- final[order(-final$Total),]
# output
sort.final
```

# Q3c Data Summary
  * The client is expecting a girl! Omit boys and give the top 10 most popular girl's names.
```{r summary3, echo=TRUE}
# filtering out the boys
filter.final <- subset(x = sort.final, sort.final$Gender != "M")
# showing top ten
top.final <- head(filter.final,10)
```

# Q3d Data Summary
  * Write these top 10 girl names and their Totals to a CSV file. Leave out the other columns entirely.
```{r summary4, echo=TRUE}
csv.final <- top.final[,c(1,5)]
write.csv(csv.final, file = "Baby_GirlNames.csv",row.names = FALSE)
```